Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-26 paper/paper.tex
comments

    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    %% WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    .htaccess access\MyComment{attempted-recon} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

\vspace{1ex}
\subsubsection{Attacks.}~Table~\ref{table:attacks} shows the selection
of attacks considered in our evaluation.  Column ``Name'' shows the
name of the attack, column ``Source'' shows the source where we
obtained data to reproduce the attack (see
Section~\ref{subsec:malicious-traffic}), column ``Type'' shows the
kind of attack, column ``Protocol'' shows the name of the protocol
explored in the attack, and column ``Ground Truth'' shows the rule
options to capture the attack. The rationale for the selection were
(1) to cover attacks with different characteristics, (2) to cover
attacks of various protocols, and (3) to explore different features of
\tname. As for (1), the selection includes attacks of the following
types: Denial-of-Service (DoS), Active Reconnaissance (AREC),
Privilege escalation (PESC), and Remote Code Execution (RCE). Note
that the distribution of attacks is not uniform across types and
protocols as these distributions are not uniform in practice. For
example, as per~\cite{emerging-threats-open}, RCE are the most
prevalent type of attacks today.\Mar{please, confirm} Columns
``Protocol'' and ``Ground Truth'' indicate, respectively, that we
analyzed attacks of different protocols and analyzed different
features of \tname.

@@

\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~For some of the attacks listed on
Table~\ref{table:attacks}, we obtained pcap files directly from
trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
input. As explained on Section~\ref{sec:minimization}, these files
abstract network messages~\cite{pcap} and can be used to avoid network
traffic during rule synthesis. In addition to these sources, we used
the Nikto2~\cite{nikto} open source security tool as follows to obtain
pcap files for http and tcp attacks. We configured an http server
(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that there are multiple instances of the same attack
encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as
  variacoes. isto eh que eh importante para a proxima sentenca.}. RQ3
evaluates the importance of these variants to \tname{}'s
performance. We used NMAP~\cite{netmap} to reproduce the Ping Scan
attack (see Section~\ref{sec:active-recon}) and hping3~\cite{hping3}
to reproduce the SYN Flood attack (see
Section~\ref{sec:dos}).

@@

\subsubsection{Benign traffic.}~Security
tools (e.g., perforation testing tools) often use benign traffic to
evaluate false positive rates. Consequently, different public options
exist. We used two popular public sources. The Bigflows.pcap data set
is part of the TcpReplay~\cite{tcpreplay} open source project. This
data set includes real network traffic of a busy private network
access point to the Internet. According to the TcpReplay web site
``This capture is much larger and has a smaller average packet size
than the previous capture (smallFlows.pcap). It also has many more
flows and different applications.''. In addition to the Bigflows.pcap
data set, we also used data sets\Mar{more than one?} made available by
Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
  of the traffic? as the description for Bigflows.pcap?}.

>>>>

    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    %% WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    .htaccess access\MyComment{attempted-recon} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

\vspace{1ex}
\subsubsection{Attacks.}~Table~\ref{table:attacks} shows the selection
of attacks considered in our evaluation.  Column ``Name'' shows the
name of the attack, column ``Source'' shows the source where we
obtained data to reproduce the attack (see
Section~\ref{subsec:malicious-traffic}), column ``Type'' shows the
kind of attack, column ``Protocol'' shows the name of the protocol
explored in the attack, and column ``Ground Truth'' shows the rule
options to capture the attack. The rationale for the selection were
(1) to cover attacks with different characteristics, (2) to cover
attacks of various protocols, and (3) to explore different features of
\tname. As for (1), the selection includes attacks of the following
types: Denial-of-Service (DoS), Active Reconnaissance (AREC),
Privilege escalation (PESC), and Remote Code Execution (RCE). Note
that the distribution of attacks is not uniform across types and
protocols as these distributions are not uniform in practice. For
example, as per~\cite{emerging-threats-open}, RCE are the most
prevalent type of attacks today.\Mar{please, confirm} \Luc{Nao da
para afirmar isso, as regras nao contem essa informacao. A classificacao
das regras eh bem generica, por exemplo: "web-application-attack".} Columns
``Protocol'' and ``Ground Truth'' indicate, respectively, that we
analyzed attacks of different protocols and analyzed different
features of \tname.

@@

\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~For some of the attacks listed on
Table~\ref{table:attacks}, we obtained pcap files directly from
trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
input. As explained on Section~\ref{sec:minimization}, these files
abstract network messages~\cite{pcap} and can be used to avoid network
traffic during rule synthesis. In addition to these sources, we used
the Nikto2~\cite{nikto} open source security tool as follows to obtain
pcap files for http and tcp attacks. We configured an http server
(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that there are multiple instances of the same attack
encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as 
variacoes. isto eh que eh importante para a proxima sentenca.}. 
\Luc{Eh feito da mesma forma q foi descrita acima. Quando um ataque 
produz varios pacotes que geram alertas, nos criamos um pcap com um 
dos pacotes e utilizamos ele como para gerar a regra maxima inicial. 
Os outros pacotes sao salvos em um outro pcap e sao utilizados para 
calcular o recall.} RQ3 evaluates the importance of these variants to \tname{}'s
performance. We used NMAP~\cite{netmap} to reproduce the Ping Scan
attack (see Section~\ref{sec:active-recon}) and hping3~\cite{hping3}
to reproduce the SYN Flood attack (see
Section~\ref{sec:dos}).

@@

\subsubsection{Benign traffic.}~Security
tools (e.g., perforation testing tools) often use benign traffic to
evaluate false positive rates. Consequently, different public options
exist. We used two popular public sources. The Bigflows.pcap data set
is part of the TcpReplay~\cite{tcpreplay} open source project. This
data set includes real network traffic of a busy private network
access point to the Internet. According to the TcpReplay web site
``This capture is much larger and has a smaller average packet size
than the previous capture (smallFlows.pcap). It also has many more
flows and different applications.''. In addition to the Bigflows.pcap
data set, we also used data sets\Mar{more than one?} made available by
Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
  of the traffic? as the description for Bigflows.pcap?}. \Luc{@Guilherme ...}
